Illuminating a Dark Lens : A Type Ia Supernova \\ Magnified by the Galaxy Cluster Abell 2744

### 2015-05-19 TABLE/magnifications2.tex
tweak to abstract; updated CATSv2.1,2.2 numbers; tweak to quality/quantity discussion

Lam(v2)        & 10 &  32 & 5 & 5  &  \citealt{Lam:2014} & Updated version of the {\tt WSLAP+} model, using more selective strong-lensing constraints and GALFIT\tablenotemark{n} models for galaxy mass.

@@

\begin{deluxetable}{p{0.62in}ccc}
\tablecolumns{4}
\tablecaption{Lens model predictions for SN \tomas\ magnification\label{tab:PredictedMagnifications}}
\tablehead{ \colhead{Model\tablenotemark{a}} & 
            \colhead{Best\tablenotemark{b}} & \colhead{Median\tablenotemark{c}} & 
            \colhead{68\% Conf. Range\tablenotemark{d}}}
\startdata
Bradac(v1)      & 3.19     & 2.48  &   2.31$-$2.66 \\
CATS(v1)        & 2.28     & 2.29  &   2.25$-$2.34 \\
CATS(v1sw)      & \nodata  & 2.62  &   2.44$-$2.80 \\
Merten          & 2.33     & 2.24  &   2.04$-$2.92 \\
Sharon(v1)      & 2.56     & 2.60  &   2.44$-$2.78 \\
Sharon(v2)      & 2.74     & 2.59  &   2.42$-$2.85 \\
Zitrin-LTM      & 2.67     & 2.99  &   2.61$-$3.77 \\
Zitrin-NFW      & 2.09     & 2.29  &   2.07$-$2.52 \\
Williams        & 2.70     & 2.81  &   1.65$-$5.54 \\[0.1em]
\tableline\\[-0.5em]
CATS(v2)        & \nodata  & 3.42  &   3.27$-$3.58 \\
Diego           &  \nodata  & 1.80  &   1.44$-$2.16 \\
GLAFIC          & 2.34     & 2.29  &   2.19$-$2.37 \\
Lam(v1)         & \nodata  & 2.79  &   2.42$-$3.16 \\[0.5em]
Bradac(v2)$^{\rm *}$ & 2.23     & 2.26  &   2.30$-$2.23 \\
Lam(v2)$^{\rm *}$ & 1.86     & 1.91  &   1.54$-$2.28
\enddata
\tablenotetext{a}{Models above the line are from the pre-HFF set, and those below incorporate HFF data. The final two, marked by an asterisk, were not part of the blind test as they include modifications made after the measured magnification of the SN was known.}  
\tablenotetext{b}{The magnification returned for the optimal version of each model, as independently defined by each lens modeling team.}  
\tablenotetext{c}{Median magnification from 100-600 Monte Carlo realizations of the model.}  
\tablenotetext{d}{Confidence ranges about the median, enclosing 68\%\ of the realized values.}
\end{deluxetable}

>>>>

Lam(v2)        & 10 &  32 & 5 & 5  &  \citealt{Lam:2014} & Updated version of the {\tt WSLAP+} model, using more selective strong-lensing constraints and GALFIT\tablenotemark{n} models for galaxy mass.\\
CATS(v2.1)     & 55 & 154 & 8 & 1  &  \citealt{Jauzac:2014c} & Updated version of the CATSv2 model, adopting the spec-z constraints used for the Bradac(v2) model.\\
CATS(v2.2)     & 25 &  72 & 8 & 1  &  \citealt{Jauzac:2014c} & Updated version of the CATSv2 model, adopting the spec-z constraints and multiple-image definitions used for the Bradac(v2) model.

@@

\begin{deluxetable}{p{0.62in}ccc}
\tablecolumns{4}
\tablecaption{Lens model predictions for SN \tomas\ magnification\label{tab:PredictedMagnifications}}
\tablehead{ \colhead{Model\tablenotemark{a}} & 
            \colhead{Best\tablenotemark{b}} & \colhead{Median\tablenotemark{c}} & 
            \colhead{68\% Conf. Range\tablenotemark{d}}}
\startdata
Bradac(v1)      & 3.19     & 2.48  &   2.31$-$2.66 \\
CATS(v1)        & 2.28     & 2.29  &   2.25$-$2.34 \\
CATS(v1sw)      & \nodata  & 2.62  &   2.44$-$2.80 \\
Merten          & 2.33     & 2.24  &   2.04$-$2.92 \\
Sharon(v1)      & 2.56     & 2.60  &   2.44$-$2.78 \\
Sharon(v2)      & 2.74     & 2.59  &   2.42$-$2.85 \\
Zitrin-LTM      & 2.67     & 2.99  &   2.61$-$3.77 \\
Zitrin-NFW      & 2.09     & 2.29  &   2.07$-$2.52 \\
Williams        & 2.70     & 2.81  &   1.65$-$5.54 \\[0.1em]
\tableline\\[-0.5em]
CATS(v2)        & \nodata  & 3.42  &   3.27$-$3.58 \\
Diego           &  \nodata  & 1.80  &   1.44$-$2.16 \\
GLAFIC          & 2.34     & 2.29  &   2.19$-$2.37 \\
Lam(v1)         & \nodata  & 2.79  &   2.42$-$3.16 \\[0.5em]
Bradac(v2)$^{\rm *}$ & 2.23     & 2.26  &   2.30$-$2.23 \\
Lam(v2)$^{\rm *}$ & 1.86     & 1.91  &   1.54$-$2.28\\
CATS(v2.1)$^{\rm *}$ & \nodata & 3.06  &   2.92$-$3.19 \\
CATS(v2.2)$^{\rm *}$ & \nodata & 3.07  &   2.94$-$3.20 \\
\enddata
\tablenotetext{a}{Models above the line are from the pre-HFF set, and those below incorporate HFF data. The final two, marked by an asterisk, were not part of the blind test as they include modifications made after the measured magnification of the SN was known.}  
\tablenotetext{b}{The magnification returned for the optimal version of each model, as independently defined by each lens modeling team.}  
\tablenotetext{c}{Median magnification from 100-600 Monte Carlo realizations of the model.}  
\tablenotetext{d}{Confidence ranges about the median, enclosing 68\%\ of the realized values.}
\end{deluxetable}

### 2015-05-19 snTomas.tex
tweak to abstract; updated CATSv2.1,2.2 numbers; tweak to quality/quantity discussion

\begin{abstract}
\changenote{Reduced to $<250$ words.}\\
SN \tomas\ is a Type Ia Supernova (SN) discovered at
$z=1.3457\pm0.0001$ behind the galaxy cluster Abell 2744
($z=0.308$). This SN 
%has a projected separation from the cluster core
%of $\sim$40\arcsec, 
is closer to the critical lensing line than any
previous cluster-lensed Type Ia SN, 
%As such, it provides 
providing a valuable
opportunity to test galaxy cluster lens models
with the direct measurement of an absolute magnification at the edge
of the strong-lensing region.  
%We derive a tightly constrained measure
%of the peak apparent magnitude, corrected for light curve shape and
%extinction.  
In a cosmology-independent analysis, we find that \tomas\ is
$0.77\pm0.15$ magnitudes brighter than unlensed Type Ia SNe at similar
redshift, implying a lensing magnification of 
$\mu_{\rm obs}=2.03\pm0.29$. 
\change{In future lens models of Abell 2744, this could
provide some leverage for breaking degeneracies that positional
constraints cannot disentangle.}  We compare the measured $\mu$
against predicted magnifications from 17 lens models, \change{ of
which 13 were generated before the SN magnification was known,
qualifying as a pure ``blind test.''}  The models are collectively
fairly accurate: 8 of the models deliver median magnifications that
are consistent with the measured $\mu$ to within 1$\sigma$.  However,
there is a subtle systematic bias: all of the models that disagree
with the SN $\mu$
%by more than $1\sigma$ 
are {\it higher} than the measured value and 6 of these
are $\geq1.5\sigma$ above the measured magnification.  We
evaluate possible causes for this mild bias,
%and reject explanations
%that invoke misinterpretation of the SN data or errors in the
%definition of the lens models. We 
and find that
% -- at least for this
%cluster and this line of sight -- 
the most accurate models are those that employ stringent quality cuts
\change{and use a large fraction of spectroscopically measured 
redshifts of multiple images.}
% A more robust evaluation
%\change{Further examination} of systematic biases 
%could be achieved 
%\change{could be pursued with more}
%lens model comparisons using lensed
%Type Ia SNe, as well as cluster lensing simulations.
\end{abstract}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{FIG/tension}
\caption{ \label{fig:Nspecz}
The model tension as a function of strong-lensing constraints. The y
axis in both panels marks the number of standard deviations in
discrepancy between the measured SN magnification and the prediction
from each lens model. The left panel plots this tension against the
total number of multiply imaged systems used, and the right panel
plots the fraction of
multiply-imaged systems that have a spectroscopic redshift
constraint.  Marker shapes and colors are the same as in
Figure~\ref{fig:LensingTest}.  \change{In both panels, three lens model sequences are highlighted with connecting lines: the Bradac, Lam, and CATS(v2) models, which are the three examples where the quantity and quality of strong lensing constraints changed substantially from one model version to the next.}
} 
\end{center}
\end{figure*}

@@

\change{
The SN \tomas\ comparison therefore supports the notion that the {\it
quality} of strong-lensing constraints is just as important as the
{\it quantity}.  To get accurate magnifications, lens models should
employ a strict vetting of the multiply-imaged systems.  Each system
used should ideally have a spectroscopic redshift, or at least a very
well constrained photometric redshift.  The primary value in this
strict vetting is to avoid erroneous redshifts being assigned to
multiply-imaged systems that are unconstrained in redshift space.
Such redshift errors are especially problematic for relatively low-$z$
background systems: at lower redshifts the model-derived deflection
angle and magnification show a more rapid variation with the redshift
of the background galaxy than for similarly positioned high redshift
galaxies.}

>>>>

\begin{abstract}
\changenote{Reduced to $<250$ words.}\\
SN \tomas\ is a Type Ia Supernova (SN) discovered at
$z=1.3457\pm0.0001$ behind the galaxy cluster Abell 2744
($z=0.308$). This SN 
%has a projected separation from the cluster core
%of $\sim$40\arcsec, 
is closer to the critical lensing line than any
previous cluster-lensed Type Ia SN, 
%As such, it provides 
providing a valuable
opportunity to test galaxy cluster lens models
with the direct measurement of an absolute magnification at the edge
of the strong-lensing region.  
%We derive a tightly constrained measure
%of the peak apparent magnitude, corrected for light curve shape and
%extinction.  
In a cosmology-independent analysis, we find that \tomas\ is
$0.77\pm0.15$ magnitudes brighter than unlensed Type Ia SNe at similar
redshift, implying a lensing magnification of 
$\mu_{\rm obs}=2.03\pm0.29$. 
\change{In future lens models of Abell 2744, this could
provide some leverage for breaking degeneracies that positional
constraints cannot disentangle.}  We compare the measured $\mu$
against predicted magnifications from 17 lens models, \change{ of
which 13 were generated before the SN magnification was known,
qualifying as a pure ``blind test.''}  The models are collectively
fairly accurate: 8 of the models deliver median magnifications that
are consistent with the measured $\mu$ to within 1$\sigma$.  However,
there is a subtle systematic bias: all of the models that disagree
with the SN $\mu$
%by more than $1\sigma$ 
are {\it higher} than the measured value and 6 of these
are $\geq1.5\sigma$ above the measured magnification.  We
evaluate possible causes for this mild bias,
%and reject explanations
%that invoke misinterpretation of the SN data or errors in the
%definition of the lens models. We 
and \changeR{find that model accuracy can be improved to some extent
by using multiply-imaged systems vetted with stringent quality cuts
and requiring a large fraction with spectroscopic redshifts.}
% -- at least for this
%cluster and this line of sight -- 
% A more robust evaluation
%\change{Further examination} of systematic biases 
%could be achieved 
%\change{could be pursued with more}
%lens model comparisons using lensed
%Type Ia SNe, as well as cluster lensing simulations.
\end{abstract}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{FIG/tension}
\caption{ \label{fig:Nspecz}
The model tension as a function of strong-lensing constraints. The y
axis in both panels marks the number of standard deviations in
discrepancy between the measured SN magnification and the prediction
from each lens model. The left panel plots this tension against the
total number of multiply imaged systems used, and the right panel
plots the fraction of
multiply-imaged systems that have a spectroscopic redshift
constraint.  Marker shapes and colors are the same as in
Figure~\ref{fig:LensingTest}.  \change{In both panels, three lens model sequences are highlighted with connecting lines: the Bradac, Lam, and CATS(v2) models, which are the three examples where the number of strong lensing constraints and the fraction with spectroscopic redshifts changed substantially from one model version to the next.}
} 
\end{center}
\end{figure*}

@@

\change{
The SN \tomas\ comparison therefore supports the notion that the {\it
quality} of strong-lensing constraints is just as important as the
{\it quantity}} \changeR{ -- at least for computing magnifications of sources outside the strong-lensing region. Once a sufficient number of robust multiply-imaged systems have been included (perhaps $\sim$10) most models will be dominated by systematic uncertainties such as those described in Section~\ref{sec:LensModelErrors}.  To improve the accuracy of magnification predictions with a large number of strong-lensing constraints, one } 
\change{
should employ a strict vetting of the multiply-imaged systems.  Each
system used should ideally have a spectroscopic redshift, or at least
a very well constrained photometric redshift.  The primary value in
this strict vetting is to avoid erroneous redshifts being assigned to
multiply-imaged systems that are unconstrained in redshift space.
Such redshift errors are especially problematic for relatively low-$z$
background systems: at lower redshifts the model-derived deflection
angle and magnification show a more rapid variation with the redshift
of the background galaxy than for similarly positioned high redshift
galaxies.}
